Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-07-26 snapshot_system/arch.tex
revised

%Our system adopts VM-centric snapshot  lazy delete strategy so that all snapshot deletions are scheduled
%in the backup time window at midnight. 
%Therefore, snapshot deletions must be fast enough to fit in time window and
%efficient enough to satisfy our resource constraints.
%However, there is no simple solution can achieve these goals with high reliability.
%Our hybrid deletion strategy, using fuzzy deletion regularly and accurate deletion periodically,
%accomplishs our speed, resource usage and relibility goals very well.
\begin{itemize}
\item {\bf Computation for snapshot fingerprint summary.}
Every time there is a new snapshot created,
we compute a bloom-filter as the summary of reference pointers for all non-PDS chunks used 
in this snapshot.
%{\bf Creating bloom filter} Scan all the living snapshot recipess and their segment recipes,
%for every reference pointing to append store, add it to the bloom filter.

\item {\bf Approximate deletion with fast summary comparison.}
Then when there is a snapshot deletion, 
we use  the snapshot fingerprint summary to identify if  chunks to be deleted from one snapshot
are still used by other snapshots. 
This is done approximately and quickly by comparing the fingerprints of deleted snapshots with
the fingerprint bloom-filter summary of other live snapshots.
Since the number of live snapshots is limited for each VM (e.g. 10 in the Alibaba's production system), 
the complexity of comparison is acceptable.
%Instead of scanning the entire append store indices, we merge the type-1 summaries of all
%valid snapshots. 
%Since each VM has uniform bloom filter parameters to create snapshot summaries, 
%such merged summeries give us a compact representation of
%all block fingerprints that are still in use.
%Thus by the property of bloom filter, if a fingerprint is not found in merged summaries, 
%we are certain that block is no longer used by any valid snapshot, it would be then added
%to append store's deletion log.
However, there is a small false-positive ratio which
would identify unused data as in use, resulting in temporary storage leakage.

@@

The cost of leak repair mainly come from holding a table of reference pointers and scan of all snapshots metadata.
Consider each reference pointer consumes 8 bytes plus 1 byte as mark field, a VM that has 40GB backup data will
need 90MB of memory to construct such a table. Also, scanning all snapshots metadata is many times slower comparing
to the approximate deletion which only scans single snapshot's metadata.
It's worth to mention that compare to previously-developed mark-and-sweep techniques, 
our leak repair still has advantage because the scope of repair is restricted within single VM's snapshot 
store due to our VM-specific design, and we only do it periodically.
%We cannot simply repeat the phase 1 multiple times to reduce temporary storage leakage, because:
%\begin{enumerate} 
%\item After several runs of phase-1, it is proven that the merged type-1 summaries cannot sieve remaining unused blocks, due to the false-positive property of bloom filter.
%\item The recipes of deleted snapshots have been removed from the system, thus we are not able to obtain the deleted block fingerprints from any metadata, the only way to discover them is to scan the append store indices.
%\end{enumerate}
\end{itemize}

@@

%{\bf Discussion}
We now give a model to estimate the storage leakage and how often leak repair needs to be conducted.
Assuming a VM always maintain $B$ snapshots in the backup, and it creates and deletes one snapshot
everyday. Let $c$ be the total number of chunks brought by the initial backup, $c^'$ be the average
number of chunks brought by a consequancial backup, then the total number of chunks in use $C$ is:
\[
C = c + (B-1)c^'
\]

Let $m$ be the bloom filter size in bits, $k$ be the number of hash functions used by bloom filter,
then bloom filter's false-positive rate $\epsilon$ is:
\[
\epsilon = (1-e^{\frac{kC}{m}})^k
\]

For each snapshot deletion, the amount of chunks need to be deleted is nearly identical to the number of
newly add chunks $c^'$. However, some of chunks among them are not detected as unused in our approximation
algorithm, thus forms the storage leakage. Let $R$ be the total number of runs of approximate deletion since
last repair, the total leakage $L$ can be estimated as:
\[
L = Rc^'\epsilon
\]

When leakage ratio $L/C$ exceeds a pre-defined threshold $\tau$, we need to execute a repare:
\[
\frac{L}{C} = \frac{Rc^'\epsilon}{c+(B-1)c^'} > \tau \rightarrow R > \frac{\tau}{\epsilon}\frac{c + (B-1)c^'}{c^'}
\]
For example, when $\epsilon = 0.01$ and $\tau=0.1$, when the VM keeps 10 snapshots and each snapshot contains
about 5\% of new data, then
we would expect leak repair be triggered once for 
every 290 runs of approximate deletion. 
When one machine hosts 25 VMs and there is one snapshot deletion per day per VM, there would be 
only one full leak repair scheduled for every 12 days, which is sufficiently small to offset the 
heavy I/O workload of the mark-sweep process.

>>>>

%Our system adopts VM-centric snapshot  lazy delete strategy so that all snapshot deletions are scheduled
%in the backup time window at midnight. 
%Therefore, snapshot deletions must be fast enough to fit in time window and
%efficient enough to satisfy our resource constraints.
%However, there is no simple solution can achieve these goals with high reliability.
%Our hybrid deletion strategy, using fuzzy deletion regularly and accurate deletion periodically,
%accomplishs our speed, resource usage and relibility goals very well.
\begin{itemize}
\item {\bf Computation for snapshot fingerprint summary.}
Every time there is a new snapshot created,
we compute a bloom-filter with $z$ bits as the summary of reference pointers for all non-PDS chunks used 
in this snapshot.

%To control the false positive ratio $\epislon$ under $0.01$, an average snapshot of size 40 $GB$ with 
%$u \approx 10$ million chunks, $z$ has 10 million bits.

%{\bf Creating bloom filter} Scan all the living snapshot recipess and their segment recipes,
%for every reference pointing to append store, add it to the bloom filter.

\item {\bf Approximate deletion with fast summary comparison.}
Then when there is a snapshot deletion,  
we need to identify if  chunks to be deleted from one snapshot
are still used by other snapshots. 
This is done approximately and quickly by comparing the reference pointers of deleted snapshots with
the merged reference bloom-filter summary of other live snapshots.
The merging of live snapshot bloom-filter bits uses the logical OR operator and the merged vector still takes $z$ bits.
Since the number of live snapshots is limited for each VM (e.g. 10 in the Alibaba's production system), 
the time cost of this comparison is small.
%Instead of scanning the entire append store indices, we merge the type-1 summaries of all
%valid snapshots. 
%Since each VM has uniform bloom filter parameters to create snapshot summaries, 
%such merged summeries give us a compact representation of
%all block fingerprints that are still in use.
%Thus by the property of bloom filter, if a fingerprint is not found in merged summaries, 
%we are certain that block is no longer used by any valid snapshot, it would be then added
%to append store's deletion log.
However, there is a small false-positive ratio which
would identify unused data as in use, resulting in temporary storage leakage.

@@

The cost of leak repair mainly come from holding a table of reference pointers and scan of all snapshots metadata.
Consider each reference pointer consumes 8 bytes plus 1 byte as mark field, a VM that has 40GB backup data with about
10 million chunks will need 90MB of memory to construct such a table. 
Also, scanning all snapshots metadata is many times slower comparing
to the approximate deletion which only scans single snapshot's metadata.
It's worth to mention that compare to previously-developed mark-and-sweep techniques, 
our leak repair still has advantage because the scope of repair is restricted within single VM's snapshot 
store due to our VM-specific design, and we only do it periodically.
%We cannot simply repeat the phase 1 multiple times to reduce temporary storage leakage, because:
%\begin{enumerate} 
%\item After several runs of phase-1, it is proven that the merged type-1 summaries cannot sieve remaining unused blocks, due to the false-positive property of bloom filter.
%\item The recipes of deleted snapshots have been removed from the system, thus we are not able to obtain the deleted block fingerprints from any metadata, the only way to discover them is to scan the append store indices.
%\end{enumerate}
\end{itemize}

@@

%{\bf Discussion}
We now estimate the storage leakage and how often leak repair needs to be conducted.
Assuming a VM always maintain $h$ snapshots in the backup, and it creates and deletes one snapshot
everyday. Let $u$ be the total number of chunks brought by the initial backup, $\Delta u$ be the average
number of additional unique chunks added from one snapshot to the next snapshot. Then the total number of unique
chunks used is about:
\[
U = u + (h-1)\Delta u.
\]

Each bloom filter vector has  $z$ bits for each snapshot and let $j$ be the number of hash functions used by the
bloom filter. The probablity that a particular bit is still in all $h$ summary vectors is still 0 is 
$(1- \frac{1}{z}) ^{j U}$. Notice that when a chunk appears multiple times in these summary vectors, it does not 
increase the probability of a particular bit being 0 in all $h$ vectors.
Then the false-positive rate $\epsilon$ is: 
\[
\epsilon = (1-(1-\frac{z})^{jU})^j.
\]

For each snapshot deletion, the amount of chunks need to be deleted is nearly identical to the number of
newly add chunks $\Delta u$. However, some of chunks among them are not detected as unused in our approximation
algorithm, thus forms the storage leakage. Let $R$ be the total number of runs of approximate deletion since
last repair. we estimate  the total leakage $L$ after $R$ runs as:
\[
L = R \Delta u \epsilon
\]

When leakage ratio $L/U$ exceeds a pre-defined threshold $\tau$, we need to execute a leak repair:
\[
\frac{L}{U} = \frac{R \Delta u \epsilon}{u+(h-1)\Delta u } > \tau 
\Longrightarrow R > \frac{\tau}{\epsilon}\frac{u + (h-1)\Delta u}{\Delta u}
\]
For example in our tested dataset,  
each VM keeps $h=10$ snapshots and each snapshot has
about 1-5\% of new data. Thus $\frac{\Delta u}{u}=0.05$ at most. With 40GB snapshot size, $u\approx = 10$ millions.
Then $U=10.45$ millions.
We choose  $\epsilon = 0.01$ and $\tau=0.1$. Then from the bloom filter  false positive formula, 
$z=10U=100.45$ million bits. $R=290$.
Then we would expect leak repair be triggered once for 
every 290 runs of approximate deletion. 
When one machine hosts 25 VMs and there is one snapshot deletion per day per VM, there would be 
only one full leak repair for one VM scheduled for every 12 days. Each repairs uses about 90MB memory on average
as discussed and takes a few minutes.
% which is sufficiently small to offset the heavy I/O workload of the mark-sweep process.

### 2013-07-26 snapshot_system/snapshot.tex
revised

\begin{itemize}
\item \textbf{Local duplicate search with VM-specific bining.}
We start with the standard dirty bit approach in a coarse grain segment level.
In our implementation with Xen at an Alibaba platform, the segment size is 2MB
and  the device driver is extended to support  changed block tracking. 
Since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in memory
and cannot afford a small segment  size.
%Xen doesn't directly support it, However, their open-source 
%architecture allows anyone to extend 
%and use the  Xen virtual device driver to implement a feature called ``changed block tracking''
%for the storage device and the dirty bit setting is maintained in a coarse grain level we call it a segment.
It should be noted that dirtybit tracking is supported or can be easily implemented in 
major virtualization solution vendors. For example,
%(VMware, Xen, Microsoft).  
%VMware support it directly; 
the VMWare hypervisor has an API to let external backup applications know 
the changed areas since last backup. 
%We implement dirty bit tracking this way in Alibaba's platform.  
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic and implement  the changed block tracking feature.

Since the best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{??},  
%This allows the system to achieve deduplication even when there are insertions/deletions which would
%affect many fixed-size chunks.
we conduct additional local inner-VM deduplication by comparing
dirty segment's chunk fingerprints to similar segments from its parent snapshot. 
For every segment, the minimum value of all its chunk fingerprints is computed during backup
and is recorded in the snapshot recipe.
So when we are processing a dirty segment,
we can easily find similar segments from the
parent snapshot recipe and load their segment recipes. Then we can perform fingerprint matching 
against them
for further inner-VM deduplication.
The amount of memory for maintaining those fingerprints is small, as we only
load one segment at a time.
For example, with a 2MB segment, there are about 500 fingerprints to compare.

>>>>

\begin{itemize}
\item \textbf{VM-specific local duplicate search within similar segments.}
We start with the standard dirty bit approach in a coarse grain segment level.
In our implementation with Xen at an Alibaba platform, the segment size is 2MB
and  the device driver is extended to support  changed block tracking. 
Since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in memory
and cannot afford a small segment  size.
%Xen doesn't directly support it, However, their open-source 
%architecture allows anyone to extend 
%and use the  Xen virtual device driver to implement a feature called ``changed block tracking''
%for the storage device and the dirty bit setting is maintained in a coarse grain level we call it a segment.
It should be noted that dirtybit tracking is supported or can be easily implemented in 
major virtualization solution vendors. For example,
%(VMware, Xen, Microsoft).  
%VMware support it directly; 
the VMWare hypervisor has an API to let external backup applications know 
the changed areas since last backup. 
%We implement dirty bit tracking this way in Alibaba's platform.  
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic and implement  the changed block tracking feature.

Since the best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{??},  
%This allows the system to achieve deduplication even when there are insertions/deletions which would
%affect many fixed-size chunks.
we conduct additional local inner-VM deduplication by comparing
dirty segment's chunk fingerprints to similar segments from its parent snapshot. 
For every segment, the minimum value of all its chunk fingerprints is computed during backup
and is recorded in the snapshot recipe.
When we are processing a dirty segment,
we can easily find similar segments from the
parent snapshot recipe and load their segment recipes. Then 
given a set of data chunks within a dirty segement,  we compare  these chunk fingerprints
with those in similar segments.  
For example, with a 2MB segment, there are about 500 fingerprints to compare.
The amount of memory for maintaining those fingerprints in similar segment s is small, as we only
load one dirty segment at a time.
